---
title: "Phase 2 - Data description"
format: html
editor: 
  markdown: 
    wrap: sentence
---

## Libraries

```{r}
#| echo: false
#| message: false
#| warning: false
library(vroom)
library(here)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
library(readr)
```

## Adding the data

```{r}
#| echo: false
#| message: false
emissions_data <- vroom(here("data", "emissions_by_unit_and_fuel_type_c_d_aa.csv"))
workforce_data <- vroom(here("data", "qwi_ca_se_f_gs_n4_oslp_u.csv"))
subsectors_type <- vroom(here("data", "subsect_types.csv"))
```


## Links and sources

### Link to the github project

https://github.com/vito100-8/Research_project.git

Due to our use of very large data files, we used Git LFS (Large File Storage) that allowed us to push our data into the project repository.

### Some infos about the data sources

We used two data sources : the US Environmental Protection Agency and the US Census Bureau.

#### Environmental Protection Agency

The Environmental Protection Agency (EPA) is an American federal agency which takes part in scientific research and enforces environmental laws by developing regulations to contribute to the preservation of the environment.
Its main missions are fighting against climate change, reducing pollution, and ensuring environmental equity.
As part of its monitoring and regulatory role, the EPA produces multiple databases, such as the Air Quality System (AQS, which contains data from air pollution monitoring stations across the U.S; Greenhouse Gas (GHG) Reporting Program which provides data on greenhouse gas emissions from large facilities.
The dataset which we will use in this project is part of the GHG Program.
All datasets of the EPA are open data, as part of the objective of scientific transparency which is promoted by the agency.
The EPA also possesses its own built-in data tools, like EnviroFacts, which allows for an immediate analysis of data.

#### United States Census Bureau

The U.S. Census Bureau, part of the Department of Commerce, is the federal agency responsible for collecting and analyzing data about the U.S. population and economy.
Its primary role is conducting the decennial U.S.
Census, which determines congressional representation, legislative district boundaries, and the allocation of federal funds for infrastructure, healthcare, etc.
The agency also conducts the Economic Census every five years, which provides data on the industries and businesses in the US.
The agency also conducts yearly, quarterly and monthly surveys about the socio-economic dynamics of the population and labor statistics.
The dataset we used is part of the Quarterly Workforce Indicators,coming from the Longitudinal Employer-Household Dynamics page.
The QWI is particular in the way that it is able to link workers to where they work, which allows for the identification of worker flows, demographics etc.
The US Census Bureau’s data is accessible and open to all.
Through the platform data.census.gov, one is able to add and filter data directly for analysis or creation of geographic maps, tables etc.

### Links to the dataset sources

Emissions by Unit and Fuel Type (from the US Environmental Protection Agency) : https://www.epa.gov/system/files/other-files/2024-10/emissions_by_unit_and_fuel_type_c_d_aa.zip

The zipped file contained an Excel binary file, with four sheets : Emissions by Unit type, Emissions by Fuel type, Industry type referencing to the first two sheets and a Q&A sheet.
We chose to focus on only the first sheet, which contained CO2 emissions of energy producers.
The third sheet containing information about the sub-sectors in which the energy facilities are listed.
We saved these two sheets in two separate CSV files.
We needed to unzip it and create a CSV manually because the native file contained lines not understandable by R.

Quarterly Workforce Indicators (QWI) Data (from the US Census Bureau) :
https://lehd.ces.census.gov/data/qwi/latest_release/al/qwi_ca_se_f_gs_n4_oslp_u.csv.gz

We manually unzipped the unique datafile before loading it into R. 

Quarterly Workforce Indicators (QWI) Data (from the US Census Bureau) : https://lehd.ces.census.gov/data/qwi/latest_release/al/qwi_al_se_fs_gs_n4_op_u.csv.gz

We manually unzipped the unique datafile before loading it into R.

## About the datasets

```{r}
#| echo: false
#Summary table for emissions
emissions_summary <- emissions_data |>
  summarise(
    "Number of rows" = nrow(emissions_data),
    "Number of columns" = ncol(emissions_data),
    "Period" = paste(range(`Reporting Year`), collapse= " - "),
    "Periodicity" = "Annual",
    "Extremum levels of CO2 emissions " = paste(range(`Unit CO2 emissions (non-biogenic)`), collapse = " - "),
    "Number of sectors" = n_distinct(unlist(strsplit(paste(`Industry Type (sectors)`, collapse = ","), ",\\s*"))), #separate the rows that have a "," to not double count firms in several subsectors as distinct values
    "Number of subparts" = n_distinct(unlist(strsplit(paste(`Industry Type (subparts)`, collapse = ","), ",\\s*"))),
    "Number of different companies" = length(unique(`Primary NAICS Code`)),
    "Total number of missing values across the sheet" = sum(sapply(emissions_data, function(x) sum(is.na(x))))
  )

emissions_summary |>
  kable("html", caption = "Emissions database information") |>
  kable_styling(full_width = T, bootstrap_options = c("striped", "hover", "condensed"))
```


```{r}
#| echo: false
#Summary table for workforce
workforce_summary <- workforce_data |>
  summarise(
    "Number of rows" = nrow(workforce_data),
    "Number of columns" = ncol(workforce_data),
    "Period" = paste(range(year), collapse= " - "),
    "Periodicity" = "Quarterly",
    "Extremums of employment (Quarterly)" = paste(range(EmpTotal, na.rm = TRUE), collapse= " - "),
    "Number of industry (NAICS 4)" = length(unique(industry)),
    "Total number of missing values across the sheet" = sum(sapply(workforce_data, function(x) sum(is.na(x))))
  )

workforce_summary |>
  kable("html", caption = "Workforce database information") |>
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "hover", "condensed"))
```

## Description of key variables

### Emissions database

`Primary NAICS Code` : this variable identifies each firm and will allow us to merge the two datasets by agregating each firm into a Naics code 4 (subsectors).
`Unit CO2 emissions (non-biogenic)`: gives the quantity of CO2 emissions for each unit (described by the variables `Unit Name` and `Unit Type`) of each firm for a given year


### Workforce database

`industry`: NAICS 4 code that identifies the subsector 
`EmpTotal`: the variable counts the estimated number of people employed in a firm at any time during the quarter according to the subsector, age and sex categories (we will agregate those specifications later on)
`EarnS` : gives us the average monthly earnings of employees with stable jobs (in dollars)



## Data treatment

Our analysis period will be 2011-2023.

### Merge preparation for the workforce dataset

```{r}
clean_wf_data <-
  workforce_data |>
  ## selection of analysis period and sector-level
  filter(year > 2010 & year < 2024) |>
  filter(ind_level == 4) |>
  ## selection of variables of interest
  select(industry, sex, education, year, quarter, EmpTotal, EarnS, FrmJbC) |>
  ## grouping to get annual instead of quarterly data
  group_by(industry, sex, education, year) |>
  summarise(
    yearly_Emp = round(mean(EmpTotal, na.rm = TRUE)),
    yearly_FrmJbC = round(mean(FrmJbC, na.rm = TRUE)),
    yearly_EarnS = mean(EarnS, na.rm = TRUE)
  )
```

### Merge preparation for the emissions dataset

```{r}
#keeping only data from 2011- 2023 
emissions_data <- emissions_data |> 
  filter(`Reporting Year` > 2010)

#keeping only the relevant columns
emissions_data <- emissions_data |>
  select(State:`Unit Type`, `Unit CO2 emissions (non-biogenic)`:`Unit Biogenic CO2 emissions (metric tons)`)

#agregating the database into a subsectors (NAICS 4) level

emissions_data <- emissions_data |> 
  filter(State == "CA") |>  # We only keep Californian firms
  mutate(
    `NAICS 4 (subsectors)` = substr(`Primary NAICS Code`, 1, 4),
    `Unit CO2 emissions (non-biogenic)` = parse_number(`Unit CO2 emissions (non-biogenic)`),  # To avoid issues     with the NA numbers during the sum after
    `Unit Methane (CH4) emissions` = parse_number(`Unit Methane (CH4) emissions`),
    `Unit Nitrous Oxide (N2O) emissions` = parse_number(`Unit Nitrous Oxide (N2O) emissions`),
  ) |> 
  group_by(`NAICS 4 (subsectors)`, `Reporting Year`, `Unit Type`) |>  # Group by NAICS subsectors and reporting year
  summarise(
    `Total CO2 Emissions (non_biogenic)` = sum(`Unit CO2 emissions (non-biogenic)`, na.rm = TRUE), 
    `Total CH4 Emissions` = sum(`Unit Methane (CH4) emissions`, na.rm = TRUE),
    `Total N20 Emissions` = sum(`Unit Nitrous Oxide (N2O) emissions`, na.rm = TRUE),
    `Number of firms` = n(),  
  )

  

```

### Merging the two datasets

```{r}
final_dataset <-
  
```

## Main target variable

## Research question



