---
title: "Phase 2 - Data description"
format: html
---

## Libraries

```{r}
#| echo: false
library(vroom)
library(here)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
```

## Adding the data

```{r}
#| echo: false
emissions_data <- vroom(here("data", "emissions_by_unit_and_fuel_type_c_d_aa.csv"))
workforce_data <- vroom(here("data", "qwi_ca_se_f_gs_n4_oslp_u.csv"))
subsectors_type <- vroom(here("data", "subsect_types.csv"))
```

## Links and sources


### Link to the github project

  https://github.com/vito100-8/Research_project.git
  
### Some infos about the data sources


### Links to the dataset sources

Emissions by Unit and Fuel Type (from the US Environmental Protection Agency) : https://www.epa.gov/system/files/other-files/2024-10/emissions_by_unit_and_fuel_type_c_d_aa.zip

THe zipped file contained an Excel binary file, with four sheets : Emissions by Unit type, Emissions by Fuel type, Industry type referencing to the first two sheets and a Q&A sheet. 
We chose to focus on only the first sheet, which contained CO2 emissions of energy producers. The third sheet containing information about the sub-sectors in which the energy facilities are listed. We saved these two sheets in two separate CSV files. 
We needed to unzip it and create a CSV manually because the native file contained lines not understandable by R.

Quarterly Workforce Indicators (QWI) Data (from the US Census Bureau) :
https://lehd.ces.census.gov/data/qwi/latest_release/al/qwi_al_se_fs_gs_n4_op_u.csv.gz

We manually unzipped the unique datafile before loading it into R. 

## About the datasets

```{r}
#| echo: false
#Summary table for emissions
emissions_summary <- emissions_data |>
  summarise(
    "Number of rows" = nrow(emissions_data),
    "Number of columns" = ncol(emissions_data),
    "Period" = paste(range(`Reporting Year`), collapse= " - "),
    "Periodicity" = "Annual",
    "Number of different companies" = length(unique(`Primary NAICS Code`)),
    "Extremum levels of CO2 emissions " = paste(range(`Unit CO2 emissions (non-biogenic)`), collapse = " - "),
    "Number of sectors" = n_distinct(unlist(strsplit(paste(`Industry Type (sectors)`, collapse = ","), ",\\s*"))), #separate the rows that have a "," to not double count firms in several subsectors as distinct values
    "Number of subparts" = n_distinct(unlist(strsplit(paste(`Industry Type (subparts)`, collapse = ","), ",\\s*"))),
    "Total number of missing values across the sheet" = sum(sapply(emissions_data, function(x) sum(is.na(x))))
    
  )


emissions_summary |>
  kable("html", caption = "Emissions database information") |>
  kable_styling(full_width = T, bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
#Summary table for workforce
workforce_summary <- workforce_data |>
  summarise(
    "Number of rows" = nrow(workforce_data),
    "Number of columns" = ncol(workforce_data),
    "Period" = paste(range(year), collapse= " - "),
    "Periodicity" = "Quarterly",
    "Number of industry (NAICS 4)" = length(unique(industry)),
    "Extremums of employment (Quarterly)" = paste(range(EmpTotal, na.rm = TRUE), collapse= " - "),
    "Total number of missing values across the sheet" = sum(sapply(workforce_data, function(x) sum(is.na(x))))
    
  )
workforce_summary |>
  kable("html", caption = "Workforce database information") |>
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "hover", "condensed"))
```

## Description of key variables

### Emissions database

### Workforce database




## Data treatment

